Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-07 paper/galsize_paper.tex
Updated draft with alternative version of color-selected clustering results, now neglecting to simultaneously address B/T and g-r, significantly simplifying the discussion with minimal loss of concrete insight

Using galaxies observed in the Sloan Digital Sky Survey (SDSS) with the stellar mass measurements taken from the MPA-JHU catalog \citep{kauffmann_etal03,brinchmann_etal04}, we define volume-limited samples of galaxies according to the same $\mstar-$completeness cuts used in \citet{behroozi_etal15} (see Figure 2). We supplement the MPA-JHU catalog with measurements of half-light radius, $\rhalf,$ derived from galaxy profile decompositions provided by \citet{meert_etal15}. The \citet{meert_etal15} catalog is based on Data Release 10 of SDSS \citep{ahn_etal14}, with improvements to the photometry pipeline and light profile fitting methods \citep{vikram_etal10,bernardi_etal13,bernardi_etal14,meert_etal13}. In the version of this catalog that we use,\footnote{Our $\mstar$ and $\rhalf$ measurements are derived from different photometry pipelines; this is driven by our desire for consistency with the \citet{behroozi_etal15} completeness cuts used in our clustering measurements. We note that we have repeated our analysis for stellar mass measurements based on \citet{meert_etl15}, finding only minor quantitative, and no qualitative changes to our results.} two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile to determine the half-light radius of total $r-$band luminosity, $\rhalf.$

@@

Using this technique, for any $\mstar-$threshold sample, the stellar mass function of the ``large" and ``small" subsamples are identical. We illustrate this for the particular case of $\mstar>10^{10.25}\msun$ in left panel of Figure~\ref{fig:sizedefinition}, which shows histograms of stellar mass for the two subsamples. The right panel of Figure~\ref{fig:sizedefinition} compares histograms of their size distributions, which partially overlap due to the variation in $\median{\rhalf}{\mstar}$ across the $\mstar-$ range of the threshold sample.

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/sdss_small_large_sample_definitions.pdf}
\caption{
{\bf Definition of ``small" and ``large" galaxies.} For a volume-limited SDSS galaxy sample defined by $M_{\ast}>10^{10.25}M_{\odot},$ we visually demonstrate how we classify galaxies into ``small" and ``large" subsamples. As described in detail in \S\ref{subsec:sizedef}, we compute the median value $\median{\rhalf}{\mstar}$ using a sliding window with a width of $1000$ galaxies at each value of $\mstar.$ The {\em left panel} compares the stellar mass functions $\phi(\mstar)$ of the two samples, confirming that our method for separating ``small" from ``large" galaxies yields subsamples with statistically indistinguishable $\phi(\mstar)$. The {\em right panel} shows histograms of $\rhalf$ for the two subsamples, which partially overlap due to the finite range of $\mstar$ in the volume-limited sample. 
}
\label{fig:sizedefinition}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

@@

In the first class of models we explore, we suppose that stellar mass $\mstar$ is the statistical regulator of $\rhalf,$ so that galaxy sizes are drawn from a log-normal distribution with $0.2$ dex of scatter centered at $\median{\rhalf}{\mstar}.$ To implement this model, for simplicity we tabulate $\median{\rhalf}{\mstar}$ directly from the data, rather than pursue a parametric form \citep[see, e.g.,][]{zhang_yang17}.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_rhalf_vs_mstar_sham_model.pdf}
\caption{
The black curve shows the median size-mass relation of SDSS galaxies as measured in \citet{meert_etal15}. The two gray bands enveloping the black curve show the $50\%$ and $90\%$ percentile regions. The blue curve shows $\rvir-$based model in which $\median{\rhalf}{\rvir}=0.01\rvir.$ This figure confirms that a linear relationship between $\rvir$ and $\rhalf,$ convolved against the nonlinear relationships between $\rvir, \mhalo$ and $\mstar,$  predicts the characteristic curvature in the relation $\median{\rhalf}{\mstar}$ over a wide range in mass.
}
\label{fig:scatter_plot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

For each volume-limited $\mstar-$threshold sample, we additionally measure $\wproj(\rproj)$ {\em without} splitting on size, giving us measurements $\wpall, \wplarge,$ and $\wpsmall$ for each threshold sample. This allows us to compute the ratio $(\wplarge-\wpsmall)/\wpall,$ which we refer to as {\em the $\rhalf$ clustering ratio}, denoted as $\delta_{\rhalf}(\wproj).$ For example, a clustering ratio of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of the same stellar mass. These ratios are the measurements appearing on the y-axis in each panel of Figure \ref{fig:clustering_ratio_upshot}.

@@

\subsection{Testing the Impact of Satellite Mass Loss}
\label{subsec:mstar_stripping}
%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=12cm]{FIGS/alt_model_wp_ratios.pdf}
\caption{
{\bf The subdominant role of tidal stripping.}
In all panels, the axes and points with error bars are the same as in Figure \ref{fig:clustering_ratio_upshot}. The solid green curves shows the prediction of the satellite mass loss model, which is in gross tension with the data due to satellites having the same size as centrals of the same mass. The solid purple curves show results for a model in which satellites lose mass after infall in a manner similar to what is seen in high-resolution hydrodynamical simulations, as described in \S\ref{subsubsec:strippingmodel}. This produces satellites that are smaller than centrals, but the effect is too mild to correctly capture the observed clustering. Evidently, satellite-specific mass stripping plays a sub-dominant role in setting the relative size of centrals and satellites.
}
\label{fig:mstarmodelclustering}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

The green curves in Figure~\ref{fig:mstarmodelclustering} test the clustering predictions of the model described in \S\ref{subsubsec:strippingmodel} {\em before} applying the satellite mass loss prescription. Thus the green curves pertain to a model where $\rhalf$ is a purely random log-normal distribution centered at $\median{\rhalf}{\mstar}.$ By construction, this model predicts no dependence at all of galaxy clustering upon $\rhalf$ at fixed $\mstar,$ simply because the scatter of $\rhalf$ about $\median{\rhalf}{\mstar}$ is uncorrelated with any other property. The application of satellite mass loss introduces correlations in the scatter: after mass loss, satellites have smaller sizes than centrals of the same $\mstar$ due to post-infall stripping.

As shown by the purple curves in Figure \ref{fig:mstarmodelclustering}, satellite mass loss impacts $\rhalf-$dependent clustering in the expected manner. The ``small" galaxy population has a higher satellite fraction due to satellites being smaller than centrals of the same $\mstar,$ and so in this model small galaxies cluster more strongly relative to large. However, the magnitude of the effect is evidently is not strong enough to produce clustering predictions that are consistent with SDSS. Evidently it is difficult to strip enough mass from satellites so that the $\rhalf-$dependent clustering is in agreement with observations. As we argue in \S\ref{sec:interpretation}, this supports the conclusion that the relative difference between central and satellite size is at least partially in place at the time of satellite infall.

\subsection{The Role of Morphology and Color}
\label{subsec:colormorph}

In this section we take a preliminary look at how galaxy clustering exhibits simultaneous dependence upon broadband color, morphology, and size. Beginning from the $\mstar>10^{10.25}\msun$ sample used above, we first divide the sample into ``red" and ``blue" subsamples according to a $g-r=0.65$ cut, the rough location of the trough of the green valley for this stellar mass. For each color-selected subsample, we separately measure $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}},$ and use these median size values to split each subsample into two.

In the top left panel of Figure \ref{fig:colorclustering} we show the clustering of large vs. small red galaxies; in the top right panel we show the same for blue galaxies. For each color-selected sample, large galaxies are slightly more strongly clustered relative to small galaxies. Comparing the top panels of Figure \ref{fig:colorclustering} to the the upper right panel of Figure \ref{fig:rvir_only_clustering_absolute} shows the dramatic impact of galaxy color selection upon $\rhalf-$dependent clustering. For $\mstar-$complete samples, small galaxies cluster much stronger than large galaxies; for color-selected samples, the reverse is true, and the magnitude of the effect weakens considerably.

The bottom panels of Figure \ref{fig:colorclustering} show results that are analogous to the top panels, but instead using of $g-r$ color, we first divide the $\mstar>10^{10.25}\msun$ sample into ``bulge-dominated" and ``disk-dominated" sequences according to $B/T,$ the fraction of $r-$band flux coming from the bulge component of the 2d light profile measurements in the \citet{meert_etal15} catalog. We define a galaxy as bulge-dominated if $B/T<0.25,$ and disk-dominated if $B/T>0.75.$ We separately split the disk-dominated and bulge-dominated subsample into two based on the median size $\median{\rhalf}{\mstar;B/T}$ appropriate for each subsample. The bottom left panel of Figure \ref{fig:colorclustering} compares the clustering of small vs. large disk-dominated galaxies; the bottom right panel shows the same comparison for bulge-dominated galaxies.

The contrast between the two bottom panels is stark. The clustering of large vs. small bulge-dominated galaxies has almost no dependence upon $\rhalf.$ On the other hand, for disk-dominated galaxies, small galaxies are much more strongly clustered, and the effect is strong. In fact, comparing the bottom left panel of Figure \ref{fig:colorclustering} to the top right panel of Figure \ref{fig:rvir_only_clustering_absolute}, we see that $\rhalf-$dependent clustering is quite similar between disk-dominated galaxies and $\mstar-$threshold samples. We discuss a halo model interpretation of these trends in \S\ref{sec:interpretation}.

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/blue_sdss_size_dependent_clustering_9p75_to_10p25.pdf}
\includegraphics[width=8cm]{FIGS/red_sdss_size_dependent_clustering_9p75_to_10p25.pdf}
\includegraphics[width=8cm]{FIGS/disk_sdss_size_dependent_clustering_9p75_to_10p25.pdf}
\includegraphics[width=8cm]{FIGS/bulge_sdss_size_dependent_clustering_9p75_to_10p25.pdf}
\caption{
{\bf Distinct $\rhalf-$dependence of clustering for color- or morphology-selected galaxy samples.}
All panels show the clustering of galaxies in the same bin of stellar mass: $10^{9.75}<\mstar/\msun<10^{10.25}.$ In the {\em upper left} panel, we first select blue galaxies based on $g-r<0.6,$ and subsequently compute $\median{\rhalf}{\mstar}$ {\em of the color-selected sample}. This results in identical stellar mass function for large and small blue galaxy samples, in analogy to the left panel of Figure \ref{fig:sizedefinition}. The {\em upper right panel} shows results of the same procedure for ``red" galaxies, defined by $g-r>0.6.$ The top two panels show that for color-selected samples, the magnitude of the $\rhalf-$dependence of clustering is dramatically reduced, and changes sign, relative to $\mstar-$complete samples. Using the \citet{meert_etal15} measurements of the disk/bulge decomposition of the 2d $r-$band luminosity profile $L_{\rm r},$ the bottom left panel shows analogous results for disk-dominated galaxies defined by $L_{\rm r}^{\rm bulge}<L_{\rm r}^{\rm tot}/4;$ the bottom right panel shows results for bulge-dominated galaxies defined by $L_{\rm r}^{\rm disk}<L_{\rm r}^{\rm tot}/4.$
}
\label{fig:colorclustering}
\end{figure*}
%---------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, we make predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the reverse is true: for both blue and red samples, $\Delta\Sigma$ of small galaxies is (weakly) suppressed relative to large galaxies. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that the analogous measurement for $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be much stronger relative to large galaxies of the same stellar mass.
}
\label{fig:lensingprediction}
\end{figure}
%-----------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_cen_sat_sizes.pdf}
\caption{
{\bf Relative sizes of centrals and satellites.}
In a narrow bin of halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun,$ we show the distribution of model galaxy sizes for different subpopulations galaxies, as predicted by the $\rvir-$based model. The red histogram shows the sizes of satellites; the blue histogram shows host halos that have never passed inside the virial radius of a larger halo (``true centrals"); the green histogram host halos that were subhalos inside a larger at some point in their past history (``splashback centrals"). In the $\rvir-$based model, galaxy size is set by the {\em physical} size of the virial radius at the time the halo attains its peak mass, naturally resulting in smaller sizes for satellites and splashback centrals relative to true centrals of the same $\mhalo.$
}
\label{fig:censatsizehist}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

The clustering measurements shown in the previous section present (at least) two intriguing puzzles. First, the modeling assumption $\rhalf\propto\rvir$ predicts size-dependent galaxy clustering that is in quite good agreement with observations, particularly considering the simplicity of the model. Yet, small galaxies cluster more strongly relative to large, which seems at odds with naive expectations based on halo mass, since $\rvir\propto\mvir.$ Second, there appears to be a highly non-trivial dependence of $\rhalf-$dependent clustering upon whether or not the galaxy sample has been selected by color or morphology. We propose a simple and unified solution to both puzzles in \S\ref{subsec:censat_sizes}, and discuss the broader implications of our interpretation in \S\ref{subsec:broader_implications}.

@@

In the $\rvir-$based model, satellite and splashback galaxies are smaller than centrals of the same halo mass due to the physical size of their halo being smaller at earlier times $\zpeak.$ There are two distinct reasons why this feature results in small galaxies being more strongly clustered relative to larger galaxies of the same mass. First and foremost, satellite galaxies statistically occupy higher mass host halos that are more strongly clustered. So in models where satellites are smaller than centrals, at fixed $\mstar$ the ``small" subsample will naturally have a higher satellite fraction, boosting the clustering of small galaxies relative to large of the same stellar mass. Second, at fixed mass, halos of $L_\ast$ galaxies that form earlier are more strongly clustered, a phenomenon commonly known as {\em halo assembly bias}. Splashback halos are typically earlier-forming than true centrals, and so models where splashback halos host smaller-than-average galaxies will naturally predict smaller galaxies being the more strongly clustered.

@@

Upon first glance, the $\rhalf-$dependence of color- and morphology-selected samples shown in Figure \ref{fig:colorclustering} appear to be driven by different principles, since the trends substantially weaken and change sign. However, even very simple models based entirely on central vs. satellite size differences can give rise to this behavior. For example, it is well known that satellites are redder than centrals \citep[e.g.,][]{vdB_etal08}; accordingly, red galaxies cluster much stronger relative to blue \citep{zehavi_etal11}. Pre-selecting galaxy samples by color will thus significantly weaken the dynamic range of $F_{\rm sat}$ within each color-selected subsample. This reduction in the role of satellites could be sufficient for the $\rhalf\propto\rvir\propto\mhalo$ of {\em central} galaxies to become the dominant effect, which would result in the mild, reversed trend seen in the top panels of Figure \ref{fig:colorclustering}. The same satellite-driven mechanism could be at work for the case of morphology pre-selection \citep[as in the morphology models developed in][]{skibba_etal08}, except that for disk-dominated systems, there would need to be a larger residual correlation with satellite designation relative to blue-selected samples \citep[which is, in fact, exactly the trend reported in][]{weinmann_etal08}.

While we consider such an explanation plausible, a conclusive determination of the principles underlying the clustering trends shown in Figure \ref{fig:colorclustering} requires proper forward modeling of morphology and color jointly with galaxy size. Such an effort is beyond the scope, but well-motivated, by the present work. See Section \ref{subsec:broader_implications} for further discussion.

@@

Recent progress has substantially improved capabilities for the richness of the predictions of empirical galaxy--halo models \citep{becker15,cohn17,moster_etal17}. This new class of models has accomplished this by empirically modeling star-formation rate, and then integrating SFR along each halo's history to differentially build up stellar mass. In ongoing follow-up work, we are currently extending this new class of models to differentially build up the disk and bulge across cosmic time using computationally efficient empirical modeling methods. In addition to treating the multi-dimensional nature of the problem, this will allow us to use data at higher redshift in a self-consistent manner.

@@

\section{Future Directions for Empirical Modeling of Galaxy Size and Morphology}
\label{sec:future}

@@

Figure \ref{fig:censatsizeratios} shows that the satellites in $\mhalo\approx10^{12}\msun$ halos can be as large as $\lesssim40-50\%$ smaller than their central galaxy counterparts. Since the satellite fraction varies quite significantly with morphology and/or color, we have demonstrated that it is entirely plausible that most of the observed environmental trends of galaxy size can be understood simply in terms of central vs. satellite size differences. While it remains to be seen whether this explanation ends up comprising the whole story, one thing is now clear: it is simply not possible to reliably analyze environmental trends of galaxy $\rhalf$ without properly accounting for satellites. Because of the large systematics that are common in contemporary group-finding methods \citep{campbell_etal15}, we consider forward modeling methods to be a requisite for meaningful continued progress on this topic.\footnote{We note that group-finders may play an important role in future analyses of galaxy size. However, it was shown in \citet{campbell_etal15} that treating observed groups as if they were correctly identified dark matter halos can lead to entirely unreliable conclusions. The robust method to extract information from galaxy group catalogs is instead to generate a Monte Carlo realization of a forward model, and then run the same group-finding algorithm on the mock as was run on the data, enabling a proper apples-to-apples comparison.}

>>>>

Using galaxies observed in the Sloan Digital Sky Survey (SDSS) with the stellar mass measurements taken from the MPA-JHU catalog \citep{kauffmann_etal03,brinchmann_etal04}, we define volume-limited samples of galaxies according to the same $\mstar-$completeness cuts used in \citet{behroozi_etal15} (see Figure 2). We supplement the MPA-JHU catalog with measurements of half-light radius, $\rhalf,$ derived from galaxy profile decompositions provided by \citet{meert_etal15}. The \citet{meert_etal15} catalog is based on Data Release 10 of SDSS \citep{ahn_etal14}, with improvements to the photometry pipeline and light profile fitting methods \citep{vikram_etal10,bernardi_etal13,bernardi_etal14,meert_etal13}. In the version of this catalog that we use,\footnote{Our $\mstar$ and $\rhalf$ measurements are derived from different photometry pipelines; this is driven by our desire for consistency with the \citet{behroozi_etal15} completeness cuts used in our clustering measurements. We note that we have repeated our analysis for stellar mass measurements based on \citet{meert_etal15}, finding only minor quantitative, and no qualitative changes to our results.} two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile to determine the half-light radius of total $r-$band luminosity, $\rhalf.$

@@

Using this technique, for any $\mstar-$threshold sample, the stellar mass function of the ``large" and ``small" subsamples are identical. We illustrate this for the particular case of $\mstar>10^{10.25}\msun$ in left panel of Figure~\ref{fig:sizedefinition}, which shows histograms of stellar mass for the two subsamples. The right panel of Figure~\ref{fig:sizedefinition} compares histograms of the two size distributions, which partially overlap due to the variation in $\median{\rhalf}{\mstar}$ across the $\mstar-$ range of the threshold sample.

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/sdss_small_large_sample_definitions.pdf}
\caption{
{\bf Definition of ``small" and ``large" galaxies.} For a volume-limited SDSS galaxy sample defined by $M_{\ast}>10^{10.25}M_{\odot},$ we visually demonstrate how we classify galaxies into ``small" and ``large" subsamples. As described in detail in \S\ref{subsec:sizedef}, we compute the median value $\median{\rhalf}{\mstar}$ at the stellar mass of each galaxy in the sample, and divide the sample into two based on this $\mstar-$dependent cut. The {\em left panel} compares the stellar mass functions $\phi(\mstar)$ of the two samples, confirming that our method for separating ``small" from ``large" galaxies yields subsamples with statistically indistinguishable $\phi(\mstar)$. The {\em right panel} shows histograms of $\rhalf$ for the two subsamples, which partially overlap due to the finite range of $\mstar$ in the volume-limited sample. 
}
\label{fig:sizedefinition}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

Figure \ref{fig:baseline_sham_clustering} shows that our implementation of SHAM produces model galaxies whose projected clustering, $\wproj(\rproj)$, is in reasonably good agreement with SDSS galaxies. We note, however, that $\wproj(\rproj\gtrsim1{\rm Mpc})$ is $\sim10-20\%$ too weak at stellar mass $\mstar\gtrsim10^{10.75}\msun.$

@@

In one class of models we explore, we suppose that stellar mass $\mstar$ is the statistical regulator of $\rhalf,$ so that galaxy sizes are drawn from a log-normal distribution with $0.2$ dex of scatter centered at $\median{\rhalf}{\mstar}.$ To implement this model, for simplicity we tabulate $\median{\rhalf}{\mstar}$ directly from the data, rather than pursue a parametric form \citep[see, e.g.,][]{zhang_yang17}.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_rhalf_vs_mstar_sham_model.pdf}
\caption{
The black curve shows the median size-mass relation of SDSS galaxies as measured in \citet{meert_etal15}. The two gray bands enveloping the black curve show the $50\%$ and $90\%$ percentile regions. The blue curve shows $\rvir-$based model in which $\median{\rhalf}{\rvir}=0.01\rvir.$ This figure confirms that a linear relationship between $\rvir$ and $\rhalf$ predicts the characteristic curvature in the relation $\median{\rhalf}{\mstar}$ over a wide range in mass.
}
\label{fig:scatter_plot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

For each volume-limited $\mstar-$threshold sample, we additionally measure $\wproj(\rproj)$ {\em without} splitting on size, giving us measurements $\wpall, \wplarge,$ and $\wpsmall$ for each threshold sample. This allows us to compute the ratio $(\wplarge-\wpsmall)/\wpall,$ which we refer to as {\em the $\rhalf$ clustering ratio}. These ratios, denoted as $\delta\wproj/\wproj,$ are the measurements appearing on the y-axis in each panel of Figure \ref{fig:clustering_ratio_upshot}. As a specific example, a clustering ratio of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of the same stellar mass.

@@

\subsection{Testing the Impact of Satellite Mass Loss}
\label{subsec:mstar_stripping}
%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=12cm]{FIGS/alt_model_wp_ratios.pdf}
\caption{
{\bf The subdominant role of tidal stripping.}
In all panels, the axes and points with error bars are the same as in Figure \ref{fig:clustering_ratio_upshot}. The solid green curves shows the prediction of a model where $\rhalf$ is statistically set by $\mstar,$ in the complete absence of satellite mass loss. Such a model would predict zero $\rhalf-$dependence to galaxy clustering, in gross tension with observations. The solid purple curves show results for a model in which satellites lose mass after infall in a manner similar to what is seen in high-resolution hydrodynamical simulations, as described in \S\ref{subsubsec:strippingmodel}. This produces satellites that are smaller than centrals, but the effect is too mild to correctly capture the observed clustering. Evidently, satellite-specific mass stripping plays a sub-dominant role in setting the relative size of centrals and satellites.
}
\label{fig:mstarmodelclustering}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

The green curves in Figure~\ref{fig:mstarmodelclustering} test the clustering predictions of the model described in \S\ref{subsubsec:strippingmodel} {\em before} applying the satellite mass loss prescription. Thus the green curves pertain to a model where $\rhalf$ follows a purely random log-normal distribution centered at $\median{\rhalf}{\mstar}.$ By construction, this model predicts no dependence at all of galaxy clustering upon $\rhalf$ at fixed $\mstar,$ simply because the scatter of $\rhalf$ about $\median{\rhalf}{\mstar}$ is uncorrelated with any other variable. The application of satellite mass loss introduces correlations in the scatter: after mass loss, satellites have smaller sizes than centrals of the same $\mstar$ due to post-infall stripping.

The purple curves in Figure \ref{fig:mstarmodelclustering} show that satellite mass loss impacts $\rhalf-$dependent clustering in the expected manner. The ``small" galaxy population has a higher satellite fraction due to satellites being smaller than centrals of the same $\mstar,$ and so in this model small galaxies cluster more strongly relative to large. However, the magnitude of the effect is not strong enough to produce clustering predictions that are consistent with SDSS. Evidently it is difficult to strip enough mass from satellites so that the $\rhalf-$dependent clustering predictions are in agreement with observations. As we argue in \S\ref{sec:interpretation}, this supports the conclusion that the relative difference between central and satellite size is at least partially in place at the time of satellite infall.

\subsection{The Role of Galaxy Color}
\label{subsec:colormorph}

In this section we take a preliminary look at how galaxy clustering exhibits simultaneous dependence upon broadband color and size. Beginning from the $\mstar>10^{10.25}\msun$ sample used above, we first divide the sample into ``red" and ``blue" subsamples according to a $g-r=0.65$ cut, the rough location of the trough of the green valley for this stellar mass. For each color-selected subsample, we separately measure $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}},$ and use these median size values to split each subsample into two.

Figure \ref{fig:colorclustering} shows the clustering of large vs. small galaxies, separately for red and blue samples. For each color-selected sample, large galaxies are slightly more strongly clustered relative to small galaxies. Comparing Figure \ref{fig:colorclustering} to the the upper right panel of Figure \ref{fig:rvir_only_clustering_absolute} shows the dramatic impact of galaxy color selection upon $\rhalf-$dependent clustering. For $\mstar-$complete samples, small galaxies cluster much stronger than large galaxies; for color-selected samples, the reverse is true, and the magnitude of the effect weakens considerably. We discuss a halo model interpretation of these trends in \S\ref{sec:interpretation}.

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/color_selected_size_dependent_clustering.pdf}
\caption{
{\bf Distinct $\rhalf-$dependence of clustering for color-selected galaxy samples.}
Projected clustering of SDSS galaxies in the bin of stellar mass: $10^{9.75}<\mstar/\msun<10^{10.25}.$ For both ``blue" and ``red" samples, we first divide SDSS galaxies falling in the $\mstar$ bin based on a color cut at $g-r=0.65.$ We then subsequently compute $\median{\rhalf}{\mstar; {\rm red}}$ and $\median{\rhalf}{\mstar; {\rm blue}}$. In analogy to the left panel of Figure \ref{fig:sizedefinition}, this sample definition results in identical stellar mass function for large and small subsamples, separately for both red and blue parent samples. We then measure $\wproj$ for each of the four subsamples, plotting the results with color-coding as specified in the legend. At this stellar mass, pre-selecting galaxy samples by color results in a dramatic weakening of the $\rhalf-$dependence of $\wproj$ relative to $\mstar-$complete samples, as well as a sign-reversal of the signal. 
}
\label{fig:colorclustering}
\end{figure*}
%---------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, we make predictions for as-yet-unseen measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show $\Delta\Sigma$ of small galaxies to be significantly stronger relative to large galaxies of the same stellar mass. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the trend is reversed and weakened: for both blue and red samples, $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. 
}
\label{fig:lensingprediction}
\end{figure}
%-----------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_cen_sat_sizes.pdf}
\caption{
{\bf Relative sizes of centrals and satellites.}
In a narrow bin of halo mass $\mhalo\equiv\mpeak\approx10^{12}\msun,$ we show the distribution of model galaxy sizes for different subpopulations galaxies, as predicted by the $\rvir-$based model. The red histogram shows the sizes of satellites; the blue histogram shows host halos that have never passed inside the virial radius of a larger halo (``true centrals"); the green histogram host halos that were subhalos inside a larger at some point in their past history (``splashback centrals"). In the $\rvir-$based model, galaxy size is set by the {\em physical} size of the virial radius at the time the halo attains its peak mass, naturally resulting in smaller sizes for satellites and splashback centrals relative to true centrals of the same $\mhalo.$
}
\label{fig:censatsizehist}
\end{figure}
%-----------------------------------------------------------------------------------------------------

@@

The clustering measurements shown in the previous section present (at least) two intriguing puzzles. First, the modeling assumption $\rhalf\propto\rvir$ predicts size-dependent galaxy clustering that is in quite good agreement with observations, particularly considering the simplicity of the model. Yet, small galaxies cluster more strongly relative to large, which seems at odds with naive expectations based on halo mass, since $\rvir\propto\mvir.$ Second, there appears to be a highly non-trivial dependence of $\rhalf-$dependent clustering upon whether or not the galaxy sample has been selected by color. We propose a simple and unified solution to both puzzles in \S\ref{subsec:censat_sizes}, and discuss the broader implications of our interpretation in \S\ref{subsec:broader_implications}.

@@

In the $\rvir-$based model, satellites and splashback galaxies are smaller than centrals of the same halo mass due to the physical size of their halo being smaller at earlier times $\zpeak.$ There are two distinct reasons why this feature results in small galaxies being more strongly clustered relative to larger galaxies of the same mass. First and foremost, satellite galaxies statistically occupy higher mass host halos that are more strongly clustered. So in models where satellites are smaller than centrals of the same mass, smaller galaxies will have a higher satellite fraction, boosting the clustering of small relative to large galaxies of the same stellar mass. Second, at fixed mass, halos of $L_\ast$ galaxies that form earlier are more strongly clustered, a phenomenon commonly known as {\em halo assembly bias} \citep{gao_white05,wechsler_etal06}. Splashback halos are typically earlier-forming than true centrals \citep{wang_etal09}, and so models where splashback halos host smaller-than-average galaxies will naturally predict smaller galaxies being the more strongly clustered \citep[see][for an example of the splashback-dependence of halo clustering]{sunayama_etal16}.

@@

Upon first glance, the $\rhalf-$dependence of color-selected samples shown in Figure \ref{fig:colorclustering} appear to be driven by different principles, since the trends substantially weaken and change sign. However, even very simple models based entirely on central vs. satellite size differences can give rise to this behavior. For example, it is well known that satellites are redder than centrals of the same stellar mass \citep[e.g.,][]{vdB_etal08}; accordingly, red galaxies cluster much stronger relative to blue \citep{zehavi_etal11}. Pre-selecting galaxy samples by color will thus significantly weaken the dynamic range of $F_{\rm sat}$ within each color-selected subsample. This reduction in the role of satellites could be sufficient for {\em central} galaxies to be responsible for the dominant effect. When the influence of centrals dominates the signal, models where $\rhalf\propto\rvir$ predict that $\rhalf-$dependent clustering should reverse sign, since $\rvir\propto\mhalo^{1/3},$ and halo bias $b(\mhalo)$ increases with halo mass. Models where $\rhalf\propto\rvir$ would also predict this reversed signal should have relatively small magnitude, since $\rvir$ scales only weakly with $\mhalo,$ and because $b(\mhalo)$ is weak function of mass for halos hosting galaxies $\mstar\lesssim10^{11}\msun.$ Therefore, if the segregation of galaxies into red- and blue-sequences is correlated sufficiently strongly with central vs. satellite designation, then the $\rhalf-$dependent clustering of color-selected samples shown in the top panels of Figure \ref{fig:colorclustering} are consistent with predictions of models where $\rhalf\propto\rvir.$

While we consider such an explanation plausible, a conclusive determination of the principles underlying the clustering trends shown in Figure \ref{fig:colorclustering} requires proper forward modeling of size and color jointly. Such an effort is beyond the scope, but well-motivated, by the present work. See Section \ref{subsec:broader_implications} for further discussion.

@@

@@

\section{Future Directions for Empirical Modeling of Galaxy Size}
\label{sec:future}

@@

Figure \ref{fig:censatsizeratios} shows that the satellites in $\mhalo\approx10^{12}\msun$ halos can be as large as $\lesssim40-50\%$ smaller than their central galaxy counterparts. Since the satellite fraction varies quite significantly with morphology and/or color, we have demonstrated that it is entirely plausible that most of the observed environmental trends of galaxy size can be understood simply in terms of central vs. satellite size differences. While it remains to be seen whether this explanation ends up comprising the whole story, one thing is now clear: it is not possible to reliably analyze environmental trends of galaxy $\rhalf$ without properly accounting for satellites. %Because of the large systematics that are common in contemporary group-finding methods \citep{campbell_etal15}, we consider forward modeling methods to be a requisite for continued progress on this topic.
